---
title: "Insurance Charge Prediction"
author: "Your Group Name Here"
format: html
---

## Introduction
This Dataset is an extensive compilation of health-related information aimed at offering insights into the factors that impact individual medical insurance expenses. This dataset covers diverse attributes related to individuals' demographic details, lifestyles, and medical backgrounds. The main emphasis is on gaining a deeper understanding of the economic aspects of healthcare, with a specific focus on the medical costs incurred by individuals.In this dataset, we have 7 variables/columns in total. 
1.Age: This is the age of the individual.
2.Sex: This refers to the gender of the individual.This categorical variable takes two values female and male. 
3.BMI (Body Mass Index): BMI is a measure of body fat based on height and weight.
4.Children: This indicates the number of children or dependents covered by the insurance. 
5.Smoker: This binary feature likely indicates whether the individual is a smoker or not. 
6.Region:This categorical variable represents the geographical region where the individual resides.There are four unique inputs, namely "southwest" "southeast" "northwest"and "northeast".
7.Charges: This is likely the target variable, representing the medical costs incurred by the individual. 
We have two categorical variables sex(with two unique inputs) and region(with four unique inputs) and one binary variable smoker.
We have 1338 rows in this dataset.
In the health insurance sector, insurance firms frequently encounter difficulties when accurately calculating premiums for individual policyholders. Inaccuracies in assessing health risks can lead to substantial financial setbacks. Therefore, precise determination of health insurance premiums is essential for ensuring the financial stability of insurance companies and delivering equitable services to policyholders. This dataset offers a substantial volume of data encompassing various facets, providing valuable insights that can assist health insurance companies in shaping their operational and business strategies.

```{r}
library(tidyverse)
library(dslabs)
library(MixAll)
library(ggfortify)
library(pheatmap)
setwd("/Users/gaoyuchen/Downloads")
data <- read.csv("insurance.csv")
ncol(data)
unique(data$sex)
unique(data$region)
unique(data$smoker)
nrow(data)

```


## Research Questions

The first hypothesis I would like to make is that individuals who smoke are likely to incur higher medical charges in comparison to those who do not smoke.
The second hypothesis I would like to make is that older individuals are more prone to incurring higher medical charges compared to their younger counterparts.
The last hypothesis is that people with higher BMI are likely to incur higher medical charges.


## Data Exploration


```{r}
boxplot(charges ~ sex, data = data,
        xlab = "sex", ylab = "charges",
        main = "Boxplot of charges by sex")


```
The boxplot above shows that sex seems to has little impact on the medical charges since the median of charges seems to be the same for female and male. However, we can see that the spread of charges value is larger for male. 

```{r}
boxplot(charges ~ smoker, data = data,
        xlab = "smoker", ylab = "charges",
        main = "Boxplot of charges by smoker")
```
The boxplot shows that people who smoke has a significantly larger charges compared with those who do not since its median is much larger.


```{r}
boxplot(charges ~ region, data = data,
        xlab = "region", ylab = "charges",
        main = "Boxplot of charges by region")
```
The plot shows that the difference in charges for people living in different regions is not very significant. People living in northeast have a little higher charges compared with other regions.

```{r}

plot(data$bmi, data$charges,
     xlab = "bmi",
     ylab = "charges",
     main = "Scatterplot of bmi vs. charges")


```
The scatterplot above shows a trend that people with higher bmi tend to have higher medical charges.Besides that, the degree of variability of the charges is higher at higher bmi which is reflected by the more scattered points.
```{r}

plot(data$age, data$charges,
     xlab = "age",
     ylab = "charges",
     main = "Scatterplot of age vs. charges")


```
The trend here is interesting. While we can observe a positive relationship between age and medical charges, the points here seem to be divided into three different groups.

```{r}
    cor <- cor(data[, sapply(data, is.numeric)])
    ## Make the heatmap
    pheatmap(cor,
             treeheight_col = 0,
             treeheight_row = 0,
             display_numbers = TRUE)

```
The heatmap above demonstrates the correlation between the numeric variables and charges. Across all the numeric variables, age has the highest correlation with the charges(0.3). From the three boxplots we got previously, we can conclude that smoker has the most significant impact on chargers, thus has higher correlation with the response variable chargers.
## Multiple Linear Regression Model

```{r}
model=lm(charges~.,data=data)
summary(model)

```
interpretation of slope for numeric variables:
The increase in age by 1 unit will lead to an increase in charges by $256.9.
The increase in bmi by 1 unit will lead to an increase in charges by $339.2.
The increase in children by 1 unit will lead to an increase in charges by $339.2.

interpretation of slope for categorical variables:
Holding other factors constant, on average the person who smoke will lead to $23848.5 increase in charges  compared with the one who does not.
On average, holding other factors constant, males are expected to have charges that are $131.3 lower compared to charges incurred by females.
Holding other factors constant, on average the persons who live in northwest region are expected to have charges that are $353 lower compared to charges incurred by people living in northeast.
Holding other factors constant, on average the persons who live in southeast region are expected to have charges that are $1035 lower compared to charges incurred by people living in northeast.
Holding other factors constant, on average the persons who live in southwest region are expected to have charges that are $960 lower compared to charges incurred by people living in northeast.
```{r}
model_null=lm(charges~1,data=data)
anova(model_null,model)

```
The null hypothesis for f test here is that all the coefficients of the numeric variables are zero, and there is no difference in charges based on sex, smoking status, and regions.
The alternative hypothesis is that there is at least one variable that has impact on the medical charges of the person.
By conducting the F test, we have found out that the p value of the model is less than 0.05, indicating that we should reject the null hypothesis and the model is significant.

```{r}
summary(model)$r.squared
```
The value of multiple r square is 0.750913.It means that 75.0913% of the variability in the dependent variable charges is accounted for by the predictor variables included in the model.


```{r}
summary(model)$sigma
```
the RSE of 6062 means that, on average, the residuals (the differences between actual and predicted values) are around 6062 units with the degree of freedom of 1329.


## Improving the Model



```{r}
autoplot(model)
```

Since the blue line on the first plot increases drastically from the middle part, we can conclude that, the errors on average is not zero. 
There is a obvious increasing trend for scale-location plot, indicating that errors do not have a constant variance. The QQ-plot shows systematic deviation from the line, showing that normality assumption is violated.
```{r}
model2=lm(sqrt(charges)~.,data=data)
autoplot(model2)
```

```{r}
library(splines)
model3=lm(charges~bmi+ns(age,2)+sex+children+smoker+region,data=data)
summary(model3)
anova(model,model3)
```
Since the p value is less than 0.05, the model with the natural spline is better.

```{r}
model4=lm(charges~children*age+bmi+sex+smoker+region,data=data)
anova(model,model4)

```
Null hypothesis: There is no interaction between children and age
Alternative Hypothesis: There is interaction between children and age
Since the p value is more than 0.05, we cannot reject the null hypothesis. We should not include interaction between children and age.

```{r}
summary(model4)
```
looking at the summary, we can see that the term children:age has a p value of 0.9094, indicating that it is not statistically significant. Thus, we should not include this interaction term in our model

## Formal Hypothesis Tests

The first hypothesis I would like to make is that individuals who smoke are likely to incur higher medical charges in comparison to those who do not smoke.
Hypothesis1:
Null:Variable Smoker has a coefficient of zero or negative number 
Alternative: The coefficient of variable Smoker is a positive number.
```{r}
summary(model)
```
From the summary table, we can see that the p value of smokeryes is < 2e-16 which is less than 0.05 and the coefficient is 23848.5 which is a positive number.Thus,we can reject the null hypothesis and conclude that individuals who smoke are likely to incur higher medical charges in comparison to those who do not smoke

The second hypothesis I would like to make is that older individuals are more prone to incurring higher medical charges compared to their younger counterparts.
Hypothesis2:
Null:Variable Age has a coefficient of zero or negative number
Alternative: The coefficient of variable Age is  a positive number.
```{r}
summary(model)
```
From the summary table, we can see that the p value of age is < 2e-16 which is less than 0.05 and the coefficient is 256.9 which is a positive number.Thus,we can reject the null hypothesis and conclude that older individuals are more prone to incurring higher medical charges compared to their younger counterparts.

The last hypothesis is that people with higher BMI are likely to incur higher medical charges.
Hypothesis3:
Null:Variable BMI has a coefficient of zero or negative number
Alternative: The coefficient of variable BMI is a positive number.
```{r}
summary(model)
```
From the summary table, we can see that the p value of age is < 2e-16 which is less than 0.05 and the coefficient is 339.2  which is a positive number.Thus,we can reject the null hypothesis and conclude that people with higher BMI are likely to incur higher medical charges.

Thus, for all three hypotheses that we posited, they are all supported by our result.
However, we only have 1338 observations in our dataset which is too little to represent the whole population. Therefore, we should also be cautious about this limitation.

```{r}
0.05/3

```
Since all the p value are still less than  0.01666667, none are insignificant after making a Bonferonni correction.The conclusion does not change

## Robustness of Results

(Your text here)

## Conclusions

(Your text here)
